Isomorphisms between algebraic datatypes

### 2025-05-26 QADT/notes.tex
new note

@@

\section{Standard ADTs}
\begin{itemize}
  \item $s(X) = 1 + 2\cdot X$, $S = \mu X. s(x)$
  \item $t(X) = 1 + X^2$, $T = \mu X. t(x)$
  \item $m(X) = 1 + X + X^2$, $M = \mu X. m(x)$
  \item $j(X) = 1 + 2X + X^2$, $J = \mu X. j(x)$
  \item $h(X) = 1 + X + X^3$, $H = \mu X. h(x)$
  \item $h(X) = 1 + X^2 + X^4$
\end{itemize}

@@

\subsection{Relating different ADTs in terms of their roots}
\begin{enumerate}
  \item $T^3 = M^2$
  \item $T^3$ and $M^2$ vs. $S$?
  \item
\end{enumerate}

>>>>

\section*{2025.05.22}
\subsection*{Misc.\ notes}
\begin{itemize}
  \item Iterating linear functors does not change the roots
  \item To compute the roots of $H(H(X))$, should solve $H(x) = r_i$, where $r_i$ are the roots of $H(x)$.
\end{itemize}

\subsection*{Questions}
\begin{enumerate}
  \item Are colimits of the diagram induced by an arbitrary polynomial functor $P$ and $P \circ P$ the same?

  (Obviously, Yes.)

  \item Can that be generalized to arbitrary unfoldings?

  Can that be generalized to ``$1.5$-unfoldings'' ?

  \item When we ``add a non-zero root'', do we automatically get a
  rig isomorphism inducing an isomorphism to the original $\mu$-type?

  \item Prove the surjectivity criterion relating the image of the fold
  from the initial algebra to the final coalgebra?

  \item Big goal: When an F-algebra structure is given by a rig iso,
  the induced fold is epi (and thus iso).

  \item What are the most generic iso rules for the mu-constructor?

  \item Can every ADT $T(\vec x)$ (with any nesting of $\mu$s) be realized
  ``with all $\mu$s on the outside''?

  Strong version:  $T(\vec x) = \mu \vec y. P(\vec x, \vec y)$, where $P(\vec x, \vec y)$ is a polynomial.

  Weak version:  $T(\vec x) = U(\vec x, \mu \vec y. P(\vec x, \vec y))$, where $U(\vec x)$,
  $P(\vec x, \vec y)$ are polynomials.

  Weaker version: Same, but allowing simultaneous solution to a system of equations
  (I.E., working with fixed points of functors on $Set^n$.)

  \item Prove that $\mu x. \mu y. P(x,y) = \mu x. P(x,x)$?

  (Almost certainly yes.)

  It would immediately follow that e.g.,
  $\mu \vec x. P(\vec x) = \mu x. P(x,\dots,x) = \mu \vec x. P(\vec{\pi x})$,
  where $\vec{\pi x}$ is any permutation/mere function
  of $\vec x$.

  \item Look at $\nat \times \nat$.

  \item Look at equational laws between rig isos/elements of the iso type.

  \item
\end{enumerate}

@@

\section{Standard ADTs}
\begin{itemize}
  \item $s(X) = 1 + 2X$, $S = \mu X. s(x)$
  \item $t(X) = 1 + X^2$, $T = \mu X. t(x)$
  \item $m(X) = 1 + X + X^2$, $M = \mu X. m(x)$
  \item $j(X) = 1 + 2X + X^2$, $J = \mu X. j(x)$
  \item $h(X) = 1 + X + X^3$, $H = \mu X. h(x)$
  \item $q(X) = 1 + X^2 + X^4$, $Q = \mu X. q(x)$
\end{itemize}

@@

\subsection{Relating different ADTs in terms of their roots}
\begin{enumerate}
  \item $T^3 = M^2$
  \item $T^3$ vs.\ $M^2$ vs.\ $S$?
  \item
\end{enumerate}
